---
title: "Thesis Code"
author: "Kiersten Barr"
date: '2022-02-09'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
setwd("~/Documents/GitHub/QMSS-Thesis")
getwd()
hsls <- read.csv("Data/hsls_17_student_pets_sr_v1_0.csv")
 
#23503 observations, 9614 variables
```

#dealing with blank data

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(scales)
library(psych)
library(magrittr)
#install.packages("stargazer")
library(stargazer)
library(Hmisc)
```

```{r}

hsls$M1MATHJOB_plotting <- case_when(hsls$M1MATHJOB == -9 ~ "Missing",
                                     hsls$M1MATHJOB == -8 ~ "Missing",
                                     hsls$M1MATHJOB == 0 ~ "No",
                                     hsls$M1MATHJOB == 1 ~ "Yes")
  
M1MATHJOB_sum <- hsls %>% 
  count(M1MATHJOB_plotting = factor(M1MATHJOB_plotting)) %>% 
  mutate(pct = prop.table(n))

M1MATHJOB_sum

M1MATHJOB_sum %>% 
  ggplot(aes(x = M1MATHJOB_plotting, y = pct, label = scales::percent(pct))) + 
  geom_col(fill = "cornflowerblue")+
  geom_text(label = with(M1MATHJOB_sum, paste0(round(pct*100,1),"% ",paste0('(', n, ')'))),
            position = position_dodge(width = .9),    # move to center of bars
            vjust = -0.5,    # nudge above top of bar
            size = 3) + 
  scale_y_continuous(labels = scales::percent)+
  labs(x='Previous Math-related Job', y='Percent of Respondents',
     title="Percentage of Students with 9th-grade Math Teachers with \nMath-related Jobs Prior to Teaching",
     subtitle = "n = 23,503"
    )
```

# Look at N1SCIJOB
```{r}
hsls$N1SCIJOB_plotting <- case_when(hsls$N1SCIJOB == -9 ~ "Missing",
                                     hsls$N1SCIJOB == -8 ~ "Missing",
                                     hsls$N1SCIJOB == 0 ~ "No",
                                     hsls$N1SCIJOB == 1 ~ "Yes")

N1SCIJOB_sum <- hsls %>% 
  count(N1SCIJOB_plotting = factor(N1SCIJOB_plotting)) %>% 
  mutate(pct = prop.table(n))

N1SCIJOB_sum %>% 
  ggplot(aes(x = N1SCIJOB_plotting, y = pct, label = scales::percent(pct))) + 
  geom_col(fill = "cornflowerblue")+
  geom_text(label = with(N1SCIJOB_sum, paste0(round(pct*100,1),"% ",paste0('(', n, ')'))),
            position = position_dodge(width = .9),    # move to center of bars
            vjust = -0.5,    # nudge above top of bar
            size = 3) + 
  scale_y_continuous(labels = scales::percent)+
  labs(x='Previous Science-Related Job', y='Percent of Students',
       title="Percentage of Students with 9th-grade SCience Teachers with \nScience-related Jobs Prior to Teaching")
```


## combined into STEM field

```{r}
hsls$STEM_teach <- case_when((hsls$N1SCIJOB_plotting == "Missing" & 
                              hsls$M1MATHJOB_plotting == "Missing") ~ "Missing",
                            (hsls$N1SCIJOB_plotting == "Yes" | hsls$M1MATHJOB_plotting == "Yes") ~ "Yes",
                            (hsls$M1MATHJOB_plotting == "No" & hsls$N1SCIJOB_plotting == "No") ~ "No",
                            (hsls$M1MATHJOB_plotting == "No" & hsls$N1SCIJOB_plotting == "Missing") ~ "No",
                            (hsls$M1MATHJOB_plotting == "Missing" & hsls$N1SCIJOB_plotting == "No") ~ "No")

STEM_sum <- hsls %>% 
  count(STEM_teach_plotting = factor(STEM_teach)) %>% 
  mutate(pct = prop.table(n))

STEM_sum %>% 
  ggplot(aes(x = STEM_teach_plotting, y = pct, label = scales::percent(pct))) + 
  geom_col(fill = "cornflowerblue")+
  geom_text(label = with(STEM_sum, paste0(round(pct*100,1),"% ",paste0('(', n, ')'))),
            position = position_dodge(width = .9),    # move to center of bars
            vjust = -0.5,    # nudge above top of bar
            size = 3) + 
  scale_y_continuous(labels = scales::percent)+
  labs(x='Previous STEM-Related Job', y='Percent of Students',
       title="Percent of Students with 9th grade STEM Teachers with Prior STEM-related Jobs",
       subtitle = "Prior to data cleanup, n = 23,503")

```


# Subset dataset to applicable variables

```{r}

#pull all variables for all regressions
vars <- 
  c("M1MATHJOB","N1SCIJOB", "STEM_teach", "N1SCIJOB_plotting", "M1MATHJOB_plotting" #teacher previous stem job
          ,"X4OCC1STEM1","X4OCCFBMRST1","X5STEMCRED", 'X4OCC30STEM1'  #entrance into STEM
          ,"X3TGPASTEM","X3THIMATH","X3THISCI",'X3TCREDSTEM','X1TXMTH','X2TXMTH' #'X1TXMTSCOR','X2TXMTSCOR' 
          , 'X3TCREDMAT', 'X3TCREDSCI', 'X3TGPAMAT', 'X3TGPASCI' #achieve in STEM
          ,"X1SEX","X1RACE", "X1SES","M1ALTCERT" ,"N1ALTCERT","M1MTHYRS912","N1SCIYRS912" #control
          , 'M1SEX','N1SEX'
)

#subsetting the data for the variables above
hsls1 <- hsls[,vars]

#rename variables
colnames(hsls1) <- 
  c("MATH_job","SCI_job", "STEM_job","SCI_job_plotting","MATH_job_plotting"
          ,"STEM_first_job","STEM_job_2016","STEM_degree",'STEM_job_30'
          , "HS_STEM_GPA","Highest_HS_Math",'Highest_HS_Sci',"Tot_HS_STEM_Cred","BY_Math_Score",'F1_Math_Score'
          , 'HS_Math_Cred', 'HS_Sci_Cred', 'HS_Math_GPA', 'HS_Sci_GPA' #achieve in STEM
          ,"student_male","student_race","student_ses","MATH_altcert","SCI_altcert","Math_Years_Taught","Sci_Years_Taught"
          , "math_teacher_sex","science_teacher_sex")


# changing -9 and -8 to NA across the dataset
hsls1 <- hsls1 %>% na_if(-9) %>% na_if(-8) %>% na_if(-7)  %>% na_if(-6)

#summary(hsls1)
```


```{r}
#investigate Enter STEM variable
hsls1$ENTERstem_prep = case_when(hsls1$STEM_first_job == 1 ~ 'Yes'
                                ,hsls1$STEM_first_job == 4~ 'Yes'
                                ,hsls1$STEM_first_job == 5~ 'Yes'
                                ,hsls1$STEM_job_2016 == 1~ 'Yes'
                                ,hsls1$STEM_job_2016 == 4~ 'Yes'
                                ,hsls1$STEM_job_2016 == 5~ 'Yes'
                                ,hsls1$STEM_job_30 == 1~ 'Yes'
                                ,hsls1$STEM_job_30 == 4~ 'Yes'
                                ,hsls1$STEM_job_30 == 5~ 'Yes'
                                ,hsls1$STEM_degree == 1~ 'Yes'
                                ,(is.na(hsls1$STEM_first_job) & is.na(hsls1$STEM_job_2016) & is.na(hsls1$STEM_degree)) ~'Missing'
                                , TRUE ~ 'No')

ENTERstem_prep <- hsls1 %>% 
  count(ENTERstem_prep = factor(ENTERstem_prep)) %>% 
  mutate(pct = prop.table(n))

ENTERstem_prep %>% 
  ggplot(aes(x = ENTERstem_prep, y = pct, label = scales::percent(pct))) + 
  geom_col(fill = "cornflowerblue")+
  geom_text(label = with(ENTERstem_prep, paste0(round(pct*100,1),"% ",paste0('(', n, ')'))),
            position = position_dodge(width = .9),    # move to center of bars
            vjust = -0.5,    # nudge above top of bar
            size = 3) + 
  scale_y_continuous(labels = scales::percent)+
  labs(x='Entered STEM Fields', y='Percent of Students',
       title="Percent of Students who Entered STEM Fields (Job or Major)",
       subtitle = "Prior to data cleanup, n = 23,503")


```


#imputing
```{r}

#missing values imputing
#GPA

hsls1$imputed_HS_STEM_GPA <- with(hsls1, impute(HS_STEM_GPA, mean))
hsls1$imputed_HS_Sci_GPA <- with(hsls1, impute(HS_Sci_GPA, mean))
hsls1$imputed_HS_Math_GPA <- with(hsls1, impute(HS_Math_GPA, mean))
hsls1$imputed_Highest_HS_Math <- with(hsls1, impute(Highest_HS_Math, mean))
hsls1$imputed_Highest_HS_Sci <- with(hsls1, impute(Highest_HS_Sci, mean))
hsls1$imputed_Tot_HS_STEM_Cred <- with(hsls1, impute(Tot_HS_STEM_Cred, mean))
hsls1$imputed_HS_Sci_Cred <- with(hsls1, impute(HS_Sci_Cred, mean))
hsls1$imputed_HS_Math_Cred <- with(hsls1, impute(HS_Math_Cred, mean))
hsls1$imputed_student_ses <- with(hsls1, impute(student_ses, mean))
hsls1$imputed_BY_Math_Score <- with(hsls1, impute(BY_Math_Score, mean))
hsls1$imputed_F1_Math_Score <- with(hsls1, impute(F1_Math_Score, mean))
hsls1$imputed_Math_Years_Taught <- with(hsls1, impute(Math_Years_Taught, mean))
hsls1$imputed_Sci_Years_Taught <- with(hsls1, impute(Sci_Years_Taught, mean))

# calculate difference in math scores
hsls1$Math_Score_Change <- hsls1$imputed_F1_Math_Score - hsls1$imputed_BY_Math_Score
```

```{r}
#prepping some fields prior to removing missing fields

#add ENTERstem field
hsls1$ENTERstem = case_when(hsls1$STEM_first_job == 1 ~ 1
                                , hsls1$STEM_first_job == 4~ 1
                                ,hsls1$STEM_first_job == 5~ 1
                                ,hsls1$STEM_job_2016 == 1~ 1
                                ,hsls1$STEM_job_2016 == 4~ 1
                                ,hsls1$STEM_job_2016 == 5~ 1
                                ,hsls1$STEM_degree == 1~ 1
                                ,hsls1$STEM_job_30 == 1~ 1
                                ,hsls1$STEM_job_30 == 4~ 1
                                ,hsls1$STEM_job_30 == 5~ 1
                                , TRUE ~ 0)


#clean up student_male field
hsls1$student_male = case_when(hsls1$student_male ==1 ~1, TRUE ~ 0)

#clean up student race field
hsls1$student_race <- 
  case_when(hsls1$student_race == -9 ~ 8,
               hsls1$student_race == -8 ~ 8,
               hsls1$student_race == 1 ~ 7,
               hsls1$student_race == 2 ~ 2,
               hsls1$student_race == 3 ~ 3,
               hsls1$student_race == 4 ~ 4,
               hsls1$student_race == 8 ~ 1,
               hsls1$student_race == 5 ~ 4,
               hsls1$student_race == 6 ~ 5,
               hsls1$student_race == 7 ~ 6,
               hsls1$student_race == 8 ~ 1,
              TRUE ~ 8)

#turn race into a factor
hsls1$student_race.fact <- factor(hsls1$student_race, levels = c(1,2,3,4,5,6,7,8)
                                , labels = c("White","Asian", "Black" , "Hispanic"
                                             , "2+ races", "Native_Haw/PI", "Amer_Ind", "Unknown"))
```


#remove missing STEM teach and others to create STEM -> STEM dataset

```{r}
dim(hsls1)
# remove missing STEM_job
hsls_subset_STEM <- subset(hsls1, STEM_job != "Missing")
dim(hsls_subset_STEM)

# only keep those indicating yes or no entry into STEM
hsls_subset_STEM <- subset(hsls_subset_STEM, (STEM_first_job >= 0 | STEM_job_2016 >=0 | STEM_degree >= 0 | STEM_job_30 >= 0))

# remove missing certification statuses
hsls_subset_STEM<- subset(hsls_subset_STEM, not(is.na(MATH_altcert)))
dim(hsls_subset_STEM)

hsls_subset_STEM<- subset(hsls_subset_STEM, not(is.na(SCI_altcert))) 
dim(hsls_subset_STEM)

dim(hsls_subset_STEM)
```


# setting up Math -> STEM achieve
```{r}

# remove missing STEM_job
dim(hsls1)
hsls_subset_MATH <- subset(hsls1, not(is.na(MATH_job)))
dim(hsls_subset_MATH)
# only keep those indicating yes or no entry into STEM
hsls_subset_MATH <- subset(hsls_subset_MATH, (STEM_first_job >= 0 | STEM_job_2016 >=0 | STEM_degree >= 0 | STEM_job_30 >= 0))

dim(hsls_subset_MATH)
# remove missing certification statuses
hsls_subset_MATH<- subset(hsls_subset_MATH, not(is.na(MATH_altcert))) 

dim(hsls_subset_MATH)

```

# Setting up Sci -> Sci Achieve
```{r}
# remove missing STEM_job
dim(hsls1)
hsls_subset_SCI <- subset(hsls1, not(is.na(SCI_job)))
dim(hsls_subset_SCI)
# only keep those indicating yes or no entry into STEM
hsls_subset_SCI <- subset(hsls_subset_SCI, (STEM_first_job >= 0 | STEM_job_2016 >=0 | STEM_degree >= 0 | STEM_job_30 >= 0))

dim(hsls_subset_SCI)
# remove missing certification statuses
hsls_subset_SCI<- subset(hsls_subset_SCI, not(is.na(SCI_altcert))) 

dim(hsls_subset_SCI)
```


```{r}
STEM_sum_1 <- hsls_subset_STEM %>% 
  count(STEM_job_plotting = factor(STEM_job)) %>% 
  mutate(pct = prop.table(n))

STEM_sum_1 %>% 
  ggplot(aes(x = STEM_job_plotting, y = pct, label = scales::percent(pct))) + 
  geom_col(fill = "cornflowerblue")+
  geom_text(label = with(STEM_sum_1, paste0(round(pct*100,1),"% ",paste0('(', n, ')'))),
            position = position_dodge(width = .9),    # move to center of bars
            vjust = -0.5,    # nudge above top of bar
            size = 3) + 
  scale_y_continuous(labels = scales::percent)+
  labs(x='Previous STEM-Related Job', y='Percent of Students',
       title="Percent of Students with 9th-grade STEM Teachers with Prior STEM-related Jobs"
       , subtitle = "After data cleanup, n = 9,960")


```

```{r}
ENTERstem <- hsls_subset_STEM %>% 
  count(ENTERstem = factor(ENTERstem)) %>% 
  mutate(pct = prop.table(n))

ENTERstem %>% 
  ggplot(aes(x = ENTERstem, y = pct, label = scales::percent(pct))) + 
  geom_col(fill = "cornflowerblue")+
  geom_text(label = with(ENTERstem, paste0(round(pct*100,1),"% ",paste0('(', n, ')'))),
            position = position_dodge(width = .9),    # move to center of bars
            vjust = -0.5,    # nudge above top of bar
            size = 3) + 
  scale_y_continuous(labels = scales::percent)+
  labs(x='Entered STEM Fields', y='Percent of Students',
       title="Percent of Students who Entered STEM Fields (Job or Major)",
       subtitle = "After data cleanup, n = 9,960")
```


```{r}
dim(hsls_subset_STEM)
dim(hsls_subset_MATH)
dim(hsls_subset_SCI)
```

# Adding Achieve STEM # to data
```{r}
hsls_subset_STEM$AchieveSTEM <- rowMeans(scale(hsls_subset_STEM[, 
  c("imputed_HS_STEM_GPA","imputed_Highest_HS_Math",'imputed_Highest_HS_Sci',"imputed_Tot_HS_STEM_Cred",'imputed_F1_Math_Score')]))

hist(hsls_subset_STEM$AchieveSTEM)

summary(alpha(hsls_subset_STEM[, c("imputed_HS_STEM_GPA","imputed_Highest_HS_Math",'imputed_Highest_HS_Sci',"imputed_Tot_HS_STEM_Cred","imputed_F1_Math_Score")]))

```

#creating achieve math
```{r}

hsls_subset_MATH$AchieveMATH <- rowMeans(scale(hsls_subset_MATH[, 
  c("imputed_HS_Math_GPA","imputed_Highest_HS_Math"
    ,"imputed_HS_Math_Cred",'imputed_F1_Math_Score')]))

hist(hsls_subset_MATH$AchieveMATH)

summary(alpha(hsls_subset_MATH[,c("imputed_HS_Math_GPA","imputed_Highest_HS_Math"
    ,"imputed_HS_Math_Cred",'imputed_F1_Math_Score')]))

```

#creating achieve science
```{r}

hsls_subset_SCI$AchieveSCI <- rowMeans(scale(hsls_subset_SCI[, 
  c("imputed_HS_Sci_GPA","imputed_Highest_HS_Sci"
    ,"imputed_HS_Sci_Cred")]))

hist(hsls_subset_SCI$AchieveSCI)

summary(alpha(hsls_subset_SCI[,c("imputed_HS_Sci_GPA","imputed_Highest_HS_Sci"
    ,"imputed_HS_Sci_Cred")]))

```

```{r}
#creating new variables, changing some.
#hsls_subset1$STEMJOB = ifelse(hsls_subset1$M1MATHJOB ==1 | hsls_subset1$N1SCIJOB == 1, 1, 0)
# hsls_subset_STEMteacher_entry$ENTERstem = case_when(hsls_subset_STEMteacher_entry$STEM_first_job == 1 ~ 1
#                                 , hsls_subset_STEMteacher_entry$STEM_first_job == 4~ 1
#                                 ,hsls_subset_STEMteacher_entry$STEM_first_job == 5~ 1
#                                 ,hsls_subset_STEMteacher_entry$STEM_job_2016 == 1~ 1
#                                 ,hsls_subset_STEMteacher_entry$STEM_job_2016 == 4~ 1
#                                 ,hsls_subset_STEMteacher_entry$STEM_job_2016 == 5~ 1
#                                 ,hsls_subset_STEMteacher_entry$STEM_degree == 1~ 1
#                                 , TRUE ~ 0)
# hsls_subset_STEMteacher_entry$student_race <- 
#   case_when(hsls_subset_STEMteacher_entry$student_race == -9 ~ 8,
#                hsls_subset_STEMteacher_entry$student_race == -8 ~ 8,
#                hsls_subset_STEMteacher_entry$student_race == 1 ~ 7,
#                hsls_subset_STEMteacher_entry$student_race == 2 ~ 2,
#                hsls_subset_STEMteacher_entry$student_race == 3 ~ 3,
#                hsls_subset_STEMteacher_entry$student_race == 4 ~ 4,
#                hsls_subset_STEMteacher_entry$student_race == 8 ~ 1,
#                hsls_subset_STEMteacher_entry$student_race == 5 ~ 4,
#                hsls_subset_STEMteacher_entry$student_race == 6 ~ 5,
#                hsls_subset_STEMteacher_entry$student_race == 7 ~ 6,
#                hsls_subset_STEMteacher_entry$student_race == 8 ~ 1,
#               TRUE ~ 8)
# 
# 
# 
# hsls_subset_STEMteacher_entry$student_race.fact <- factor(hsls_subset_STEMteacher_entry$student_race, levels = c(1,2,3,4,5,6,7,8)
#                                 , labels = c("White","Asian", "Black" , "Hispanic"
#                                              , "2+ races", "Native_Haw/PI", "Amer_Ind", "Unknown"))
# 
# counts <- table(hsls_subset_STEMteacher_entry$student_race)
# #counts
# counts <- table(hsls_subset_STEMteacher_entry$student_race)
# #counts
# counts <- table(hsls_subset_STEMteacher_entry$student_race.fact)
# #counts
# barplot(counts)
```



```{r}
race_sum <- hsls_subset_STEM %>% 
  count(race = factor(student_race.fact)) %>% 
  mutate(pct = prop.table(n))

race_sum %>% 
  ggplot(aes(x = race, y = pct, label = scales::percent(pct))) + 
  geom_col(fill = "cornflowerblue")+
  geom_text(label = with(race_sum, paste0(round(pct*100,1),"% ",paste0('(', n, ')'))),
            position = position_dodge(width = .9),    # move to center of bars
            vjust = -0.5,    # nudge above top of bar
            size = 3) + 
  scale_y_continuous(labels = scales::percent)+
  labs(x='Ethnicity', y='Percent of Students',
       title="Percent of Students with 9thgrade STEM Teachers with Prior STEM-related Jobs")

```
# Descriptive Stats - Imputation comparisons

```{r}
#Dataset A: math
pre_imputation <- 
   c("HS_Math_GPA","Highest_HS_Math","HS_Math_Cred","Math_Years_Taught","F1_Math_Score", "student_ses")

post_imputation <- c("imputed_HS_Math_GPA", "imputed_Highest_HS_Math","imputed_HS_Math_Cred","imputed_Math_Years_Taught","imputed_F1_Math_Score","imputed_student_ses")

#subsetting the data for the variables above
pre_imputed_hsls_subset_MATH <- hsls_subset_MATH[,pre_imputation]

post_imputed_hsls_subset_MATH <- hsls_subset_MATH[,post_imputation]

stargazer(pre_imputed_hsls_subset_MATH, title="Pre-imputation", align=TRUE, type = "html", out = "~/Documents/GitHub/QMSS-Thesis/math_pre_imputation.html", no.space=TRUE, median= TRUE, min.max = FALSE)

stargazer(post_imputed_hsls_subset_MATH, title="Post_imputation", align=TRUE, type = "html", out = "~/Documents/GitHub/QMSS-Thesis/math_post_imputation.html", no.space=TRUE, median= TRUE, min.max = FALSE)
```

```{r}
#Dataset B: Science
pre_imputation <- 
   c("HS_Sci_GPA","Highest_HS_Sci","HS_Sci_Cred","Sci_Years_Taught","student_ses")

post_imputation <- c("imputed_HS_Sci_GPA", "imputed_Highest_HS_Sci","imputed_HS_Sci_Cred","imputed_Sci_Years_Taught","imputed_student_ses")

#subsetting the data for the variables above
pre_imputed_hsls_subset_SCI <- hsls_subset_SCI[,pre_imputation]

post_imputed_hsls_subset_SCI <- hsls_subset_SCI[,post_imputation]

stargazer(pre_imputed_hsls_subset_SCI, title="Pre-imputation", align=TRUE, type = "html", out = "~/Documents/GitHub/QMSS-Thesis/sci_pre_imputation.html", no.space=TRUE, median= TRUE, min.max = FALSE)

stargazer(post_imputed_hsls_subset_SCI, title="Post_imputation", align=TRUE, type = "html", out = "~/Documents/GitHub/QMSS-Thesis/sci_post_imputation.html", no.space=TRUE, median= TRUE, min.max = FALSE)

```


```{r}
#Dataset C: STEM
pre_imputation <- 
   c("HS_Math_GPA","HS_Sci_GPA","Highest_HS_Math","Highest_HS_Sci","Tot_HS_STEM_Cred"
     ,"Math_Years_Taught","Sci_Years_Taught","F1_Math_Score","student_ses")

post_imputation <-   
  c("imputed_HS_Math_GPA","imputed_HS_Sci_GPA","imputed_Highest_HS_Math","imputed_Highest_HS_Sci","imputed_Tot_HS_STEM_Cred"
     ,"imputed_Math_Years_Taught","imputed_Sci_Years_Taught","imputed_F1_Math_Score","imputed_student_ses")

#subsetting the data for the variables above
pre_imputed_hsls_subset_STEM <- hsls_subset_STEM[,pre_imputation]

post_imputed_hsls_subset_STEM <- hsls_subset_STEM[,post_imputation]

stargazer(pre_imputed_hsls_subset_STEM, title="Pre-imputation", align=TRUE, type = "html", out = "~/Documents/GitHub/QMSS-Thesis/stem_pre_imputation.html", no.space=TRUE, median= TRUE, min.max = FALSE)

stargazer(post_imputed_hsls_subset_STEM, title="Post_imputation", align=TRUE, type = "html", out = "~/Documents/GitHub/QMSS-Thesis/stem_post_imputation.html", no.space=TRUE, median= TRUE, min.max = FALSE)
```


## Descriptive Stats - ALL
```{r}
names(hsls_subset_STEM)

```


```{r}
#Dataset A: math
fields <- 
   c("MATH_job","AchieveMATH","ENTERstem","student_male","student_race.fact"
     ,"imputed_student_ses", "MATH_altcert","imputed_Math_Years_Taught")

#subsetting the data for the variables above
descripive_hsls_subset_MATH <- hsls_subset_MATH[,fields]

stargazer(descripive_hsls_subset_MATH, title="Pre-imputation", align=TRUE, type = "html", out = "~/Documents/GitHub/QMSS-Thesis/math_descriptive stats.html", no.space=TRUE, median= TRUE)

```
```{r}
#Dataset B: sci
fields <- 
   c("SCI_job","AchieveSCI","ENTERstem","student_male","student_race.fact"
     ,"imputed_student_ses", "SCI_altcert","imputed_Sci_Years_Taught")

#subsetting the data for the variables above
descriptive_hsls_subset_SCI <- hsls_subset_SCI[,fields]
 
stargazer(descriptive_hsls_subset_SCI, title="Descriptive stats", align=TRUE, type = "html", out = "~/Documents/GitHub/QMSS-Thesis/sci_descriptive stats.html", no.space=TRUE, median= TRUE)




```


```{r}

hsls_subset_STEM$STEM_job <- case_when(hsls_subset_STEM$STEM_job == "Yes" ~1,
                                       hsls_subset_STEM$STEM_job == "No" ~0,
                                       TRUE ~ 2)

#Dataset C: stem
fields <- 
   c("STEM_job","AchieveSTEM","ENTERstem","student_male","student_race.fact"
     ,"imputed_student_ses","MATH_altcert" ,"SCI_altcert","imputed_Math_Years_Taught"
     ,"imputed_Sci_Years_Taught")

#subsetting the data for the variables above
descriptive_hsls_subset_STEM <- hsls_subset_STEM[,fields]
 
stargazer(descriptive_hsls_subset_STEM, title="Descriptive stats", align=TRUE, type = "html", out = "~/Documents/GitHub/QMSS-Thesis/STEM_descriptive stats.html", no.space=TRUE, median= TRUE)





```

# MODELS

## Affects on Achievement


### Model 2: OLS: MATH JOB -> AchieveMATH
```{r}

lm_math = lm(AchieveMATH ~ MATH_job + student_male + student_race.fact + imputed_student_ses+MATH_altcert+imputed_Math_Years_Taught, data=hsls_subset_MATH)

summary(lm_math)
```


### Model 3: OLS: SCI JOB -> AchieveSCI

```{r}
lm_sci = lm(AchieveSCI ~ SCI_job + student_male + student_race.fact + imputed_student_ses+SCI_altcert+imputed_Sci_Years_Taught, data=hsls_subset_SCI)

summary(lm_sci)
```


### Model 1: OLS: STEM JOB -> AchieveSTEM
```{r}

lm_stem = lm(AchieveSTEM ~ STEM_job + student_male + student_race.fact + imputed_student_ses+MATH_altcert+SCI_altcert+imputed_Math_Years_Taught+imputed_Sci_Years_Taught, data=hsls_subset_STEM)

summary(lm_stem)
```

```{r}

stargazer(lm_math, lm_sci, lm_stem, title="Results", align=TRUE, type = "html", out = "~/Documents/GitHub/QMSS-Thesis/ols_results.html", no.space=TRUE)

```


#Hypothesis 2: Affect on Entrance into STEM

#MATH
```{r}
logit1_math = glm(ENTERstem ~ MATH_job+ student_male + student_race.fact + imputed_student_ses+MATH_altcert+imputed_Math_Years_Taught, data=hsls_subset_MATH, family=binomial)
summary(logit1_math)

```

```{r}
exp(coef(logit1_math))
```

#Science 
```{r}
logit2_sci = glm(ENTERstem ~ SCI_job+ student_male + student_race.fact + imputed_student_ses+SCI_altcert+ imputed_Sci_Years_Taught, data=hsls_subset_SCI, family=binomial)
summary(logit2_sci)

```

```{r}
exp(coef(logit2_sci))
```


#STEM 
```{r}
logit3_stem = glm(ENTERstem ~ STEM_job+ student_male + student_race.fact + imputed_student_ses+MATH_altcert+SCI_altcert+imputed_Math_Years_Taught+imputed_Sci_Years_Taught, data=hsls_subset_STEM, family=binomial)
summary(logit3_stem)

```

```{r}
exp(coef(logit3_stem))
```

```{r}
stargazer(logit1_math, logit2_sci, logit3_stem, title="Results", align=TRUE, type = "html", out = "~/Documents/GitHub/QMSS-Thesis/logit_models.html", no.space=TRUE)
```

```{r}

stargazer(logit1_math, logit2_sci, logit3_stem, apply.coef = exp, title="Results", align=TRUE,
          report = "vc*", single.row = TRUE, 
          se = NULL, type = "html", out = "~/Documents/GitHub/QMSS-Thesis/odds_ratios.html", no.space=TRUE)
```

# Possibilites for Appendix
##  attempting males / females

```{r}
hsls_subset_STEMteacher_entry$STEM_teacher_female = ifelse(hsls_subset_STEMteacher_entry$math_teacher_sex == 2 | hsls_subset_STEMteacher_entry$science_teacher_sex == 2, 1, 0)

hsls_subset_STEMteacher_entry_female <- subset(hsls_subset_STEMteacher_entry, student_male == 1)# & STEM_teacher_female == 1)

logit1 = glm(ENTERstem ~ STEM_job + student_race.fact + student_ses+MATH_altcert+SCI_altcert+Math_Years_Taught+Sci_Years_Taught, data=hsls_subset_STEMteacher_entry_female, family=binomial)
summary(logit1)

```

```{r}
hsls_subset_STEMteacher_entry$STEM_teacher_female = ifelse(hsls_subset_STEMteacher_entry$math_teacher_sex == 2 | hsls_subset_STEMteacher_entry$science_teacher_sex == 2, 1, 0)

#hsls_subset_STEMteacher_entry

lm3 = lm(AchieveSTEM ~ STEM_job + student_male + student_race.fact + imputed_student_ses+MATH_altcert+SCI_altcert+imputed_Math_Years_Taught+imputed_Sci_Years_Taught, data=hsls_subset_STEMteacher_entry) ## predicting how far students go in math from their math self-efficacy
summary(lm3)
```